### 2016-11-08 paper1.tex
Fixed labels

In this paper, we begin by explaining the theoretical underpinning and
numerical implementation of the superbubble-based feedback method in section~\ref{method}. 
In section~\ref{cluster} we use a single star cluster to illustrate the
effectiveness of our model at capturing the basic behaviour of superbubbles at
high and low resolution.  In section~\ref{galaxy} we apply the model to
simulations of isolated galaxies to explore the impact on the galaxy
scale ISM and the production of outflows.

\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/onestar_column.eps}
    \caption[Column density plot of a superbubble from a single cluster]{Column
    density projections from the simulations of a single star cluster with mass
    $3\times10^4\;\mathrm{M_\odot}$, $50\;\mathrm{Myr}$ after the cluster has
    formed.  The superbubble feedback model (center column) produces bubbles
    with radius and enclosed mass that match well to the direct injection
    simulations (left column).  This is shown quantitatively in
    figure~\ref{onestar_homogeneous} and figure~\ref{onestar_clumpy}.  The
    simple model (right column), despite injecting the same amount of energy,
    fails to generate enough hot mass inside the bubble, and subsequently
    suppresses the growth of Vishniac instabilities along the bubble edge
    because of the poorly resolved hot interior.}
    \label{onestar_column}
\end{figure*}

\section{Thermal Conduction and Feedback Method}\label{method}
Our new treatment of feedback has three components.  The first is the addition
of thermal conduction.   Inside resolved hot bubbles, thermal conduction
maintains uniform temperatures.   In the presence of strong gradients, thermal
conduction can lead to evaporative mass flows from cold to hot gas.  The second
component is a stochastic model of evaporation to allow resolved hot gas to
continue to gain mass from nearby cold gas.  Thus, the amount of cold gas heated
by feedback is not a free parameter, but is set by thermal conduction.  
Without a mechanism like thermal conduction, this hot gas mass is set by
how many fluid elements have feedback energy deposited into them.   It is
important to note that in our model these processes operate everywhere
temperatures are above $10^5K$.

@@

\subsection{Thermal Conduction}
In an ionized gas, thermal conduction, mediated by electrons,
transports heat down temperature gradients.  This flux, $\mathbf{Q} =
-\kappa\nabla T$, depends on the temperature gradient and the conduction
coefficient, as derived by \citet{Cowie1977},
\begin{equation} \label{coefficient}
    \kappa(T) = 1.8\left(\frac{2}{\pi}\right)^{3/2}\frac{T^{5/2}k_B^{7/2}}{m_e\,e^4\ln\Lambda}.
\end{equation}
This coefficient depends only on the Coulomb logarithm $\ln\Lambda$ (which has
an extremely weak dependence on density), and is well approximated by $\kappa(T)
= \kappa_0\,T^{5/2}$, where $\kappa_0$ is $6.1\times10^{-7}\;\mathrm{erg\;s^{-1}
K^{-7/2}cm^{-1}}$ in the absence of magnetic fields.  In order for this
situation to not cause spontaneous currents, a corresponding mass flux,
{\it in the opposite direction}, must also occur \citep{Cowie1977}.  With
spherical symmetry, this implies a mass flow rate that depends on the sound
speed in the hot gas, $c_{\rm s}$, as follows,
\begin{equation}\label{conservation}
    \frac{5}{2}\dot M c_{\rm s}^2 = 4\,\pi^2r^2\kappa(T)\frac{{\rm d}T}{{\rm d}r}.
\end{equation}

These rates hold only when the mean free path of electrons in the medium is
smaller than the scale length of the temperature gradient.  If the gradient
becomes steep enough, the heat flux (and corresponding mass flux) saturates at a
value that depends only on the density, temperature, and thermal velocity of the
electrons in the medium \citep{Cowie1977},
\begin{equation}\label{heatflux}
    \mathbf{Q} = \nabla\left(\frac{3}{2}n_ek_BT_ev_e\right).
\end{equation}
This saturation has the convenient numerical side effect of setting the smallest
timestep required to resolve this mass flux to $\sim 1/17$ the standard Courant
time.

@@

\subsection{Evaporation}
The dominant physical process governing mass flux between the hot and cold
phases of a feedback bubble is thermal conduction between the dense shell and
the hot interior.   As this process takes place on length scales far below the
resolution limit, we must capture its effects in a subgrid model.  For the thin
shell surrounding a feedback bubble, thermal conduction causes an evaporative
mass flux from the cold shell into the hot bubble.  Following
\citet{MacLow1988}, the mass flux into a bubble with interior temperature $T$
is,
\begin{equation}\label{massflux}
    \frac{{\rm d }M_b}{{\rm d}t} = \frac{4\,\pi\mu}{25k_B}\kappa_0\,\frac{\Delta T^{5/2}}{\Delta x}\ A,
\end{equation}
where $ A$ is the bubble surface area and $\Delta x$ is the thickness of
the hot layer.  For the Smoothed Particle Hydrodynamics (SPH) method used for
our tests, we calculate evaporation using the outer layer of hot particles
bordering the cold gas.  The members of this layer are those with no other hot
particles within $45^o$ of the vector to the centre of mass of their cold
neighbours.  As we cannot directly determine the radius of a well-resolved
superbubble without expensive non-local calculations, we need a way to allow
each particle to estimate the radius and thus its fractional contribution of the
shell surface area.  The total area estimated by all hot particles in a bubble
should approach $\sim 4\pi R^2$ where $R$ is the bubble radius.  For a poorly
resolved bubble $R \sim 1-2\,h$, where $h$ is a hot particle's SPH smoothing
length.  For larger bubbles each hot particle contributes an area of $\sim h^2$
and each particle sees a nearly plane-parallel section of the cold shell.  We
examined a number of bubbles at different stages of growth, as well as a
plane-parallel slab, and empirically found that a good per-particle area
estimate was $A = \frac{6\,\pi\,h^2}{N_{\rm hot}}$, where $N_{\rm
hot}$ is the number of hot neighbours for that particle and $\Delta x = h$.  We
stress that this is a fit for our specific SPH neighbour approach (described
below) that should be recalibrated for other codes.

The mass evaporation rate can be converted into a probability that a resolution
element with cold mass $m$ converts into a hot one over a time
period $\Delta t$ as follows, 
\begin{equation}\label{evaporation_probability}
    P_{evap} = \frac{{\rm d}M_b}{{\rm d}t} \frac{\Delta t}{m}.
\end{equation}
This allows us to stochastically choose full particles to evaporate, and prevent
overcooling due to fractional particle evaporation (exactly the same overcooling
problem seen in feedback heating).  Each hot particle determines how many
cold-shell particles $N_{evap}$ will evaporate each timestep, and then chooses
the $N_{evap}$ nearest cold particles, and averages the thermal energy of itself
and those particles.  These particles spontaneously join the hot bubble
(demonstrated in the figures in section~\ref{clumpy}).  The thermal conduction
rates calculated are capped using the saturation values derived in
\citet{Cowie1977}.  Within an SPH framework, this is well approximated using,
\begin{equation}\label{saturation_rate}
    \frac{{\rm d}M_{sat}}{{\rm d}t} = 17\rho\,c_{\rm s} h^2.
\end{equation}

\begin{figure*}
    \includegraphics[width=0.8\textwidth]{figures1/onestar_homogeneous.eps}
    \caption[Feedback effects as a function of resolution]{Feedback effects as a
    function of resolution.  The above figure shows on the left the total radial
    momentum imparted to the medium, and on the right the total amount of mass
    heated to above $10^5\;\mathrm{K}$.  The red curves are the direct injection
    results, the blue corresponds to the simple feedback, while the green shows
    the results of the superbubble model.  Solid lines show the values from
    simulations run with resolutions of $128^3$ particles, dashed for $64^3$,
    and dotted for $32^3$.  Note that for the $32^3$ run, the simple model
    produces $\sim 2\times 10^5\;M_\odot$ of hot gas.}
    \label{onestar_homogeneous}
\end{figure*}

@@

\begin{equation}\label{multiphase_fractions}
    m = m_{\rm hot}+m_{\rm cold}\\
    E = u_{\rm hot}m_{\rm hot}+u_{\rm cold}m_{\rm cold}
\end{equation}

Assuming pressure equilibrium, both phases will have the same pressure $P$, and
their densities are found using this and the total density $\rho$:
\begin{equation}\label{multiphase_pressurebalance}
    \frac{P}{\rho} = \frac{(\gamma-1)E}{m}
\end{equation}
\begin{equation}\label{multiphase_densities}
    \rho_{\rm cold} = \frac{P}{(\gamma-1)u_{\rm cold}}\\
    \rho_{\rm hot} = \frac{P}{(\gamma-1)u_{\rm hot}}
\end{equation}

Both the cold and hot phases are allowed to radiatively cool using their
separate temperatures and densities.   When $PdV$ work is done to a multiphase
particle, it is shared between the two phases weighted by their respective
fraction of the total energy E,
\begin{equation}\label{multiphase_PdV}
    \dot u_{\rm PdV,cold} = m\ \dot u_{\rm PdV}\frac{u_{\rm cold}}{E}\\
\end{equation}
\begin{equation}
    \dot u_{\rm PdV,hot} = m\ \dot u_{\rm PdV}\frac{u_{\rm hot}}{E}.
\end{equation}
In the absence of heating and cooling, this allows each phase to correctly
maintain constant entropy as the densities change.

@@

\begin{equation}\label{multiphase_evaporation}
    \frac{{\rm d}M_b}{{\rm d}t} = \frac{16\pi\mu}{25k_B} \kappa_0 (T_{\rm hot}^{5/2})\ h.
\end{equation}

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{figures1/onestar_hottemp.eps}
    \caption[Peak temperature of feedback-heated gas in a superbubble]{Peak
    temperature of feedback-heated gas inside the isolated star cluster's hot
    bubble.  Evaporation quickly enters the self-regulating regime, and the
    temperature is roughly constant for the last $15\;\mathrm{Myr}$ of
    feedback.}
    \label{onestar_hottemp}
\end{figure}

\section{Simulations}
\subsection{High-Resolution Star Cluster Test}\label{cluster}
We began by exploring models of a single, isolated superbubbles at high
resolution.  For these tests, we employed three physical models.  The {\it
Direct injection} approach models as much physics as possible from first
principles.  Feedback mass and energy is modelled via a stream of new gas
particles created from the star cluster.  This approach only works when the gas
resolution elements are much smaller than the star cluster mass.  The only
component of this model that is sub-grid is evaporation as this occurs on
extremely small length scales.  {\it Superbubble Feedback} refers to our new
model.  The key addition over the direct approach is the sub-grid multiphase
treatment.  Feedback energy and mass is injected into existing particles which
may split.

@@

Figure~\ref{onestar_column} shows the column density projection for the three
different feedback models in an isotropic medium.  injection model.   The hot
interior of the bubble produced using the simple feedback model contains less
mass at $128^3$ than the direct injection model and is subsequently too poorly
resolved for the instabilities formed in the accelerated shell
\citep{Vishniac1983} to mix the bubble and shell through turbulence and
diffusion.

As figure~\ref{onestar_homogeneous} shows, the superbubble feedback model
performs much better than the simple model in reproducing the hot mass
production rate from the high-resolution direct injection model, and lacks the
extreme resolution sensitivity that the simple model exhibits for the amount of
gas heated. The simple model only heats $\sim 4$ gas particles.  For the $128^3$
and $64^3$ resolutions, this gives too little hot mass.  For $32^3$ (extremely
poor resolution, in fact with gas particles more massive than the entire
cluster), this produces more than twice too much hot mass (more than
$2\times10^5\;M_\odot$). Meanwhile the simulation with the superbubble feedback
model only underestimates the hot mass by roughly a third of the target $128^3$
direct injection simulation when used at a resolution $32^3$.  The decreased
momentum in the lower-resolution runs is likely due to particles staying longer
in the multiphase state.  More massive particles take longer to fully evaporate,
and thus some mass that would be part of a pressure-driven cold shell at higher
resolution is tied up in the cold part of multiphase particles.

Figure~\ref{onestar_hottemp}  shows that the actual peak temperature of the
feedback-heated bubble in the isolated star cluster run is roughly
$1\times10^7\mathrm{K}$.

\begin{figure}
    \includegraphics[width=0.6\textwidth]{figures1/onestar_kappa.eps}
    \caption[Dependence of superbubble heated gas on conduction coefficient]{Hot
    mass production as a function of the conduction coefficient $\kappa_0$.  As
    this figure shows, reducing $\kappa_0$ by a factor of 100 reduces the amount
    of hot mass generated through conduction by only a factor of $\sim2$. All
    $\kappa_0$ values have units of $\mathrm{erg\; s^{-1} K^{-7/2} cm^{-1}}$.}
    \label{onestar_kappa}
\end{figure}

@@

Figure~\ref{onestar_kappa} shows the hot mass generated in simulations with 3
different values of $\kappa_0$.  The dashed curve shows a reasonable lower limit
for $\kappa_0$ \citep{Cowie1977}, while the dotted curve shows a much more
extreme conduction reduction than is expected in nature.  Both curves illustrate
the insensitivity of the method to reductions in the conduction rate.  Even
reducing $\kappa_0$ by 100 only results in a reduction of the hot mass inside
the bubble to just around a third.

\begin{figure}
    \includegraphics[width=\textwidth]{figures1/onestar_clumpypoints.eps}
    \caption[Image of superbubbles in a clumpy ISM]{3 kpc wide slices
    from the same three methods shown in figure~\ref{onestar_column}, also at 50
    Myr, but applied to a star cluster in an inhomogeneous medium.  Particles
    whose smoothing length intersects z=0~kpc are shown.  Particles above $10^5$
    K are shown in red.  The upper row shows simulations with a single clump,
    while the bottom row shows simulations with 6 clumps.}
    \label{onestar_clumpyslice}
\end{figure}

\begin{figure*}
    \includegraphics[width=0.8\textwidth]{figures1/onestar_clumpy.eps}
    \caption[Feedback effects as a function of ISM homogeneity]{Feedback effects
    as a function of ISM homogeneity.  The above figure shows on the left the
    total radial momentum imparted to the medium, and on the right the total
    amount of mass heated to above $10^5\;\mathrm{K}$.  The red curves are the
    direct injection results, the blue corresponds to the simple feedback, while
    the green shows the results of the superbubble model.  Solid curves show the
    results for a homogeneous ISM, dashed for a single dense clump, and dotted
    for 6 dense clumps.}
    \label{onestar_clumpy}
\end{figure*}

\subsubsection{Clumpy Medium}\label{clumpy}
As the real ISM is highly inhomogeneous (and as \citet{Silich1996} showed, cold
clumps can also be a source of evaporated cold material) , we ran two additional
simulations with cold, dense clumps with $100\;\mathrm{cm^{-3}}$ density and
$10\;\mathrm{K}$ gas in an ambient medium of $0.5\;\mathrm{cm^{-3}}$ at
$1000\:\mathrm{K}$.  This gives roughly the same amount of mass enclosed within
the hot bubble at $30\:\mathrm{Myr}$.  The first contains a single spherical
clump with a radius of $0.2\;\mathrm{kpc}$.  The second contains the same amount
of cold mass spread over 6 clumps arranged at the center of each face of a cube
$0.2\;\mathrm{kpc}$ surrounding the star.  We use this idealized clumpy medium
to test the sensitivity of the model to small scale structure that may be
unresolved in lower resolution simulations. Figure~\ref{onestar_clumpyslice}
shows a slice showing SPH particles at $50\;\mathrm{Myr}$ for the two clumpy
cases.

Figure~\ref{onestar_clumpy} shows that the superbubble model is also capable of
handling feedback into a clumpy medium.  With more than an order of magnitude
difference in the total hot mass compared to the simple model, along with less
sensitivity to environment or resolution, the superbubble model is better
suited for probing galactic outflows in simulations.

\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/disc_phase.eps}
    \caption[Phase diagram of gas in isolated galaxies]{Phase diagrams for the
    Milky Way (left) and dwarf (right) simulations at $200\;\mathrm{Myr}$.  The
    central panel shows a typical path for gas ejected from the Milky Way.
    First gas cools radiatively (cyan path) to high density, where it becomes
    multiphase as its hot component is heated to $\sim10^7\;\mathrm{K}$ by
    feedback (red path) from nearby stars formed from its neighbouring high
    density gas.  The hot phase cools primarily through adiabatic expansion
    (green path).  This process is often repeated one or more times, with gas
    that is entirely hot phase being ejected from the disc.  Cooling times of
    $10^4\;\mathrm{yr}$ and $10^6\;\mathrm{yr}$ are shown in blue. The
    majority of the hot gas in the upper left quadrant of the phase diagram has
    cooling times $>10^8\;\mathrm{yr}$.  The cooling curve for
    $10^8\;\mathrm{yr}$ passes through the green curve.  Note that particles in
    the multiphase state show the temperature and density for each state as
    separate points.  The mean properties of multiphase particles would place
    them in regions with short cooling times.}
    \label{disc_phase}
\end{figure*}

\subsection{Galaxy Simulations}\label{galaxy}
\subsubsection{Initial Conditions and Parameters}
\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/milkyway_column.eps}
    \caption[Column density and temperature images of isolated Milky Way]{Column
    Density (upper row) and temperature for the Milky Way simulation at
    $300\;\mathrm{Myr}$.  The vectors show the in-plane velocity.  Temperatures
    shown are averages between the two phases for multiphase particles.  Black
    points in the face-on images show stars formed within the last
    $20\;\mathrm{Myr}$. Note that gas is both leaving the disc near the galactic
    core, and returning in some places near the edge.}
    \label{milkyway_column}
\end{figure*}

@@

\subsubsection{ISM properties and Star Formation Rates}\label{ISM}
\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/milkyway_outflow.eps}
    \caption[Outflow evolution in isolated Milky Way]{Outflow evolution for the
    Milky Way-like simulation.  The left hand plot shows the star formation rate
    and the outflow rate The right hand plot shows the average outflow
    velocity.}
    \label{milkyway_outflow}
\end{figure*}

>>>>

In this paper, we begin by explaining the theoretical underpinning and
numerical implementation of the superbubble-based feedback method in
section~\ref{method1}. 
In section~\ref{cluster1} we use a single star cluster to illustrate the
effectiveness of our model at capturing the basic behaviour of superbubbles at
high and low resolution.  In section~\ref{galaxy1} we apply the model to
simulations of isolated galaxies to explore the impact on the galaxy
scale ISM and the production of outflows.

\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/onestar_column.eps}
    \caption[Column density plot of a superbubble from a single cluster]{Column
    density projections from the simulations of a single star cluster with mass
    $3\times10^4\;\mathrm{M_\odot}$, $50\;\mathrm{Myr}$ after the cluster has
    formed.  The superbubble feedback model (center column) produces bubbles
    with radius and enclosed mass that match well to the direct injection
    simulations (left column).  This is shown quantitatively in
    figure~\ref{onestar_homogeneous1} and figure~\ref{onestar_clumpy1}.  The
    simple model (right column), despite injecting the same amount of energy,
    fails to generate enough hot mass inside the bubble, and subsequently
    suppresses the growth of Vishniac instabilities along the bubble edge
    because of the poorly resolved hot interior.}
    \label{onestar_column1}
\end{figure*}

\section{Thermal Conduction and Feedback Method}\label{method1}
Our new treatment of feedback has three components.  The first is the addition
of thermal conduction.   Inside resolved hot bubbles, thermal conduction
maintains uniform temperatures.   In the presence of strong gradients, thermal
conduction can lead to evaporative mass flows from cold to hot gas.  The second
component is a stochastic model of evaporation to allow resolved hot gas to
continue to gain mass from nearby cold gas.  Thus, the amount of cold gas heated
by feedback is not a free parameter, but is set by thermal conduction.  
Without a mechanism like thermal conduction, this hot gas mass is set by
how many fluid elements have feedback energy deposited into them.   It is
important to note that in our model these processes operate everywhere
temperatures are above $10^5K$.

@@

\subsection{Thermal Conduction}
In an ionized gas, thermal conduction, mediated by electrons,
transports heat down temperature gradients.  This flux, $\mathbf{Q} =
-\kappa\nabla T$, depends on the temperature gradient and the conduction
coefficient, as derived by \citet{Cowie1977},
\begin{equation} \label{coefficient1}
    \kappa(T) = 1.8\left(\frac{2}{\pi}\right)^{3/2}\frac{T^{5/2}k_B^{7/2}}{m_e\,e^4\ln\Lambda}.
\end{equation}
This coefficient depends only on the Coulomb logarithm $\ln\Lambda$ (which has
an extremely weak dependence on density), and is well approximated by $\kappa(T)
= \kappa_0\,T^{5/2}$, where $\kappa_0$ is $6.1\times10^{-7}\;\mathrm{erg\;s^{-1}
K^{-7/2}cm^{-1}}$ in the absence of magnetic fields.  In order for this
situation to not cause spontaneous currents, a corresponding mass flux,
{\it in the opposite direction}, must also occur \citep{Cowie1977}.  With
spherical symmetry, this implies a mass flow rate that depends on the sound
speed in the hot gas, $c_{\rm s}$, as follows,
\begin{equation}\label{conservation1}
    \frac{5}{2}\dot M c_{\rm s}^2 = 4\,\pi^2r^2\kappa(T)\frac{{\rm d}T}{{\rm d}r}.
\end{equation}

These rates hold only when the mean free path of electrons in the medium is
smaller than the scale length of the temperature gradient.  If the gradient
becomes steep enough, the heat flux (and corresponding mass flux) saturates at a
value that depends only on the density, temperature, and thermal velocity of the
electrons in the medium \citep{Cowie1977},
\begin{equation}\label{heatflux1}
    \mathbf{Q} = \nabla\left(\frac{3}{2}n_ek_BT_ev_e\right).
\end{equation}
This saturation has the convenient numerical side effect of setting the smallest
timestep required to resolve this mass flux to $\sim 1/17$ the standard Courant
time.

@@

\subsection{Evaporation}
The dominant physical process governing mass flux between the hot and cold
phases of a feedback bubble is thermal conduction between the dense shell and
the hot interior.   As this process takes place on length scales far below the
resolution limit, we must capture its effects in a subgrid model.  For the thin
shell surrounding a feedback bubble, thermal conduction causes an evaporative
mass flux from the cold shell into the hot bubble.  Following
\citet{MacLow1988}, the mass flux into a bubble with interior temperature $T$
is,
\begin{equation}\label{massflux1}
    \frac{{\rm d }M_b}{{\rm d}t} = \frac{4\,\pi\mu}{25k_B}\kappa_0\,\frac{\Delta T^{5/2}}{\Delta x}\ A,
\end{equation}
where $ A$ is the bubble surface area and $\Delta x$ is the thickness of
the hot layer.  For the Smoothed Particle Hydrodynamics (SPH) method used for
our tests, we calculate evaporation using the outer layer of hot particles
bordering the cold gas.  The members of this layer are those with no other hot
particles within $45^o$ of the vector to the centre of mass of their cold
neighbours.  As we cannot directly determine the radius of a well-resolved
superbubble without expensive non-local calculations, we need a way to allow
each particle to estimate the radius and thus its fractional contribution of the
shell surface area.  The total area estimated by all hot particles in a bubble
should approach $\sim 4\pi R^2$ where $R$ is the bubble radius.  For a poorly
resolved bubble $R \sim 1-2\,h$, where $h$ is a hot particle's SPH smoothing
length.  For larger bubbles each hot particle contributes an area of $\sim h^2$
and each particle sees a nearly plane-parallel section of the cold shell.  We
examined a number of bubbles at different stages of growth, as well as a
plane-parallel slab, and empirically found that a good per-particle area
estimate was $A = \frac{6\,\pi\,h^2}{N_{\rm hot}}$, where $N_{\rm
hot}$ is the number of hot neighbours for that particle and $\Delta x = h$.  We
stress that this is a fit for our specific SPH neighbour approach (described
below) that should be recalibrated for other codes.

The mass evaporation rate can be converted into a probability that a resolution
element with cold mass $m$ converts into a hot one over a time
period $\Delta t$ as follows, 
\begin{equation}\label{evaporation_probability1}
    P_{evap} = \frac{{\rm d}M_b}{{\rm d}t} \frac{\Delta t}{m}.
\end{equation}
This allows us to stochastically choose full particles to evaporate, and prevent
overcooling due to fractional particle evaporation (exactly the same overcooling
problem seen in feedback heating).  Each hot particle determines how many
cold-shell particles $N_{evap}$ will evaporate each timestep, and then chooses
the $N_{evap}$ nearest cold particles, and averages the thermal energy of itself
and those particles.  These particles spontaneously join the hot bubble
(demonstrated in the figures in section~\ref{clumpy1}).  The thermal conduction
rates calculated are capped using the saturation values derived in
\citet{Cowie1977}.  Within an SPH framework, this is well approximated using,
\begin{equation}\label{saturation_rate1}
    \frac{{\rm d}M_{sat}}{{\rm d}t} = 17\rho\,c_{\rm s} h^2.
\end{equation}

\begin{figure*}
    \includegraphics[width=0.8\textwidth]{figures1/onestar_homogeneous.eps}
    \caption[Feedback effects as a function of resolution]{Feedback effects as a
    function of resolution.  The above figure shows on the left the total radial
    momentum imparted to the medium, and on the right the total amount of mass
    heated to above $10^5\;\mathrm{K}$.  The red curves are the direct injection
    results, the blue corresponds to the simple feedback, while the green shows
    the results of the superbubble model.  Solid lines show the values from
    simulations run with resolutions of $128^3$ particles, dashed for $64^3$,
    and dotted for $32^3$.  Note that for the $32^3$ run, the simple model
    produces $\sim 2\times 10^5\;M_\odot$ of hot gas.}
    \label{onestar_homogeneous1}
\end{figure*}

@@

\begin{equation}\label{multiphase_fractions1}
    m = m_{\rm hot}+m_{\rm cold}\\
    E = u_{\rm hot}m_{\rm hot}+u_{\rm cold}m_{\rm cold}
\end{equation}

Assuming pressure equilibrium, both phases will have the same pressure $P$, and
their densities are found using this and the total density $\rho$:
\begin{equation}\label{multiphase_pressurebalance1}
    \frac{P}{\rho} = \frac{(\gamma-1)E}{m}
\end{equation}
\begin{equation}\label{multiphase_densities1}
    \rho_{\rm cold} = \frac{P}{(\gamma-1)u_{\rm cold}}\\
    \rho_{\rm hot} = \frac{P}{(\gamma-1)u_{\rm hot}}
\end{equation}

Both the cold and hot phases are allowed to radiatively cool using their
separate temperatures and densities.   When $PdV$ work is done to a multiphase
particle, it is shared between the two phases weighted by their respective
fraction of the total energy E,
\begin{equation}\label{multiphase_PdV1}
    \dot u_{\rm PdV,cold} = m\ \dot u_{\rm PdV}\frac{u_{\rm cold}}{E}\\
\end{equation}
\begin{equation}
    \dot u_{\rm PdV,hot} = m\ \dot u_{\rm PdV}\frac{u_{\rm hot}}{E}.
\end{equation}
In the absence of heating and cooling, this allows each phase to correctly
maintain constant entropy as the densities change.

@@

\begin{equation}\label{multiphase_evaporation1}
    \frac{{\rm d}M_b}{{\rm d}t} = \frac{16\pi\mu}{25k_B} \kappa_0 (T_{\rm hot}^{5/2})\ h.
\end{equation}

@@

\begin{figure}
    \includegraphics[width=0.5\textwidth]{figures1/onestar_hottemp.eps}
    \caption[Peak temperature of feedback-heated gas in a superbubble]{Peak
    temperature of feedback-heated gas inside the isolated star cluster's hot
    bubble.  Evaporation quickly enters the self-regulating regime, and the
    temperature is roughly constant for the last $15\;\mathrm{Myr}$ of
    feedback.}
    \label{onestar_hottemp1}
\end{figure}

\section{Simulations}
\subsection{High-Resolution Star Cluster Test}\label{cluster1}
We began by exploring models of a single, isolated superbubbles at high
resolution.  For these tests, we employed three physical models.  The {\it
Direct injection} approach models as much physics as possible from first
principles.  Feedback mass and energy is modelled via a stream of new gas
particles created from the star cluster.  This approach only works when the gas
resolution elements are much smaller than the star cluster mass.  The only
component of this model that is sub-grid is evaporation as this occurs on
extremely small length scales.  {\it Superbubble Feedback} refers to our new
model.  The key addition over the direct approach is the sub-grid multiphase
treatment.  Feedback energy and mass is injected into existing particles which
may split.

@@

Figure~\ref{onestar_column1} shows the column density projection for the three
different feedback models in an isotropic medium.  injection model.   The hot
interior of the bubble produced using the simple feedback model contains less
mass at $128^3$ than the direct injection model and is subsequently too poorly
resolved for the instabilities formed in the accelerated shell
\citep{Vishniac1983} to mix the bubble and shell through turbulence and
diffusion.

As figure~\ref{onestar_homogeneous1} shows, the superbubble feedback model
performs much better than the simple model in reproducing the hot mass
production rate from the high-resolution direct injection model, and lacks the
extreme resolution sensitivity that the simple model exhibits for the amount of
gas heated. The simple model only heats $\sim 4$ gas particles.  For the $128^3$
and $64^3$ resolutions, this gives too little hot mass.  For $32^3$ (extremely
poor resolution, in fact with gas particles more massive than the entire
cluster), this produces more than twice too much hot mass (more than
$2\times10^5\;M_\odot$). Meanwhile the simulation with the superbubble feedback
model only underestimates the hot mass by roughly a third of the target $128^3$
direct injection simulation when used at a resolution $32^3$.  The decreased
momentum in the lower-resolution runs is likely due to particles staying longer
in the multiphase state.  More massive particles take longer to fully evaporate,
and thus some mass that would be part of a pressure-driven cold shell at higher
resolution is tied up in the cold part of multiphase particles.

Figure~\ref{onestar_hottemp1}  shows that the actual peak temperature of the
feedback-heated bubble in the isolated star cluster run is roughly
$1\times10^7\mathrm{K}$.

\begin{figure}
    \includegraphics[width=0.6\textwidth]{figures1/onestar_kappa.eps}
    \caption[Dependence of superbubble heated gas on conduction coefficient]{Hot
    mass production as a function of the conduction coefficient $\kappa_0$.  As
    this figure shows, reducing $\kappa_0$ by a factor of 100 reduces the amount
    of hot mass generated through conduction by only a factor of $\sim2$. All
    $\kappa_0$ values have units of $\mathrm{erg\; s^{-1} K^{-7/2} cm^{-1}}$.}
    \label{onestar_kappa1}
\end{figure}

@@

Figure~\ref{onestar_kappa1} shows the hot mass generated in simulations with 3
different values of $\kappa_0$.  The dashed curve shows a reasonable lower limit
for $\kappa_0$ \citep{Cowie1977}, while the dotted curve shows a much more
extreme conduction reduction than is expected in nature.  Both curves illustrate
the insensitivity of the method to reductions in the conduction rate.  Even
reducing $\kappa_0$ by 100 only results in a reduction of the hot mass inside
the bubble to just around a third.

\begin{figure}
    \includegraphics[width=\textwidth]{figures1/onestar_clumpypoints.eps}
    \caption[Image of superbubbles in a clumpy ISM]{3 kpc wide slices
    from the same three methods shown in figure~\ref{onestar_column1}, also at 50
    Myr, but applied to a star cluster in an inhomogeneous medium.  Particles
    whose smoothing length intersects z=0~kpc are shown.  Particles above $10^5$
    K are shown in red.  The upper row shows simulations with a single clump,
    while the bottom row shows simulations with 6 clumps.}
    \label{onestar_clumpyslice1}
\end{figure}

\begin{figure*}
    \includegraphics[width=0.8\textwidth]{figures1/onestar_clumpy.eps}
    \caption[Feedback effects as a function of ISM homogeneity]{Feedback effects
    as a function of ISM homogeneity.  The above figure shows on the left the
    total radial momentum imparted to the medium, and on the right the total
    amount of mass heated to above $10^5\;\mathrm{K}$.  The red curves are the
    direct injection results, the blue corresponds to the simple feedback, while
    the green shows the results of the superbubble model.  Solid curves show the
    results for a homogeneous ISM, dashed for a single dense clump, and dotted
    for 6 dense clumps.}
    \label{onestar_clumpy1}
\end{figure*}

\subsubsection{Clumpy Medium}\label{clumpy1}
As the real ISM is highly inhomogeneous (and as \citet{Silich1996} showed, cold
clumps can also be a source of evaporated cold material) , we ran two additional
simulations with cold, dense clumps with $100\;\mathrm{cm^{-3}}$ density and
$10\;\mathrm{K}$ gas in an ambient medium of $0.5\;\mathrm{cm^{-3}}$ at
$1000\:\mathrm{K}$.  This gives roughly the same amount of mass enclosed within
the hot bubble at $30\:\mathrm{Myr}$.  The first contains a single spherical
clump with a radius of $0.2\;\mathrm{kpc}$.  The second contains the same amount
of cold mass spread over 6 clumps arranged at the center of each face of a cube
$0.2\;\mathrm{kpc}$ surrounding the star.  We use this idealized clumpy medium
to test the sensitivity of the model to small scale structure that may be
unresolved in lower resolution simulations. Figure~\ref{onestar_clumpyslice1}
shows a slice showing SPH particles at $50\;\mathrm{Myr}$ for the two clumpy
cases.

Figure~\ref{onestar_clumpy1} shows that the superbubble model is also capable of
handling feedback into a clumpy medium.  With more than an order of magnitude
difference in the total hot mass compared to the simple model, along with less
sensitivity to environment or resolution, the superbubble model is better
suited for probing galactic outflows in simulations.

\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/disc_phase.eps}
    \caption[Phase diagram of gas in isolated galaxies]{Phase diagrams for the
    Milky Way (left) and dwarf (right) simulations at $200\;\mathrm{Myr}$.  The
    central panel shows a typical path for gas ejected from the Milky Way.
    First gas cools radiatively (cyan path) to high density, where it becomes
    multiphase as its hot component is heated to $\sim10^7\;\mathrm{K}$ by
    feedback (red path) from nearby stars formed from its neighbouring high
    density gas.  The hot phase cools primarily through adiabatic expansion
    (green path).  This process is often repeated one or more times, with gas
    that is entirely hot phase being ejected from the disc.  Cooling times of
    $10^4\;\mathrm{yr}$ and $10^6\;\mathrm{yr}$ are shown in blue. The
    majority of the hot gas in the upper left quadrant of the phase diagram has
    cooling times $>10^8\;\mathrm{yr}$.  The cooling curve for
    $10^8\;\mathrm{yr}$ passes through the green curve.  Note that particles in
    the multiphase state show the temperature and density for each state as
    separate points.  The mean properties of multiphase particles would place
    them in regions with short cooling times.}
    \label{disc_phase1}
\end{figure*}

\subsection{Galaxy Simulations}\label{galaxy1}
\subsubsection{Initial Conditions and Parameters}
\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/milkyway_column.eps}
    \caption[Column density and temperature images of isolated Milky Way]{Column
    Density (upper row) and temperature for the Milky Way simulation at
    $300\;\mathrm{Myr}$.  The vectors show the in-plane velocity.  Temperatures
    shown are averages between the two phases for multiphase particles.  Black
    points in the face-on images show stars formed within the last
    $20\;\mathrm{Myr}$. Note that gas is both leaving the disc near the galactic
    core, and returning in some places near the edge.}
    \label{milkyway_column1}
\end{figure*}

@@

\subsubsection{ISM properties and Star Formation Rates}\label{ISM1}
\begin{figure*}
    \includegraphics[width=\textwidth]{figures1/milkyway_outflow.eps}
    \caption[Outflow evolution in isolated Milky Way]{Outflow evolution for the
    Milky Way-like simulation.  The left hand plot shows the star formation rate
    and the outflow rate The right hand plot shows the average outflow
    velocity.}
    \label{milkyway_outflow1}
\end{figure*}
